### 2016-03-28 Rebuttal2Murali.txt

1. Why waste time describing LTSes:
We introduced LTSes mainly to define the notion of refinement that we are
using, as well as to define the operators to combine LTSes.

2. The user of a theorem-prover has to walk the "protocol space" through case
analysis:
The theorem-prover has to walk the protocol space through case analysis, but
the proof-script automation in Coq significantly reduces the burden. One can
write proof-scripts that performs a proof-search, i.e. tries to apply
previously proven basic lemmas on each of the cases and leaves the user with
very few actual proof obligations that are complex or unique so as to not be
discharged using automated proof-search.

3. Invariants are the hardest part in parametric verification of a cache
coherence protocol, and we do not mention it:
We do mention the high level theorems that we use in the paper.  While we
restrict ourselves to the verification of the MSI protocol's implementation,
the proof is modular, and a large portion of the proof can be used, as is, for
verification of other protocols (like MESI, MOSI and MOESI), as we discuss in
the previous rebuttal. The theorem prover helps construct and prove the high
level theorems that make these protocols correct. These theorems depend on a
few basic properties that can be established easily, but were not mentioned in
the paper due to lack of space. These properties include a) state changes
occuring when and only when response messages are sent, b) absence of response
messages from the parent to a child and from the same child to the parent
simultaneously in the network, for the same address and c) FIFO properties of
the messages. These basic properties seem to be the foundation over which the
other cache coherence protocols lie -- they eliminate the corner cases which
make the proof difficult.  For example, since responses from a child to its
parent and from the parent to the same child for the same address cannot be
present simultaneously, an upgrade response from the parent cannot cross a
downgrade response from the child, which would have eventually led to the
upgrade of the child and downgrade of the parent violating Lemma 3. As pointed
out by Reviewer-E, the theorems and these properties serve as a method to
derive a correct protocol (the different protocols are different system
instances which obey these properties).

4. Pipelining protocol engines and Bluespec:
Though the semantics of the atomic transitions in our paper execute one at a
time, different independent transitions can be made to occur in parallel in the
actual implementation without changing the semantics; Bluespec already does
this optimization. One can take this further, where each atomic transition can
be spread across multiple hardware clock cycles, while ensuring that the
overall atomic transaction semantics are obeyed. Bluespec currently does not
implement this, though there has been work in this area [Karczmarek et.al.].

5. Nirav's point: I don't see that Reviewer-E's scathing concern fall
into the core of what we're addressing; our focus on modularity is to
allow verification to be (mostly) carried over to different
microarchitectural variants, which is a bare minimum if one wants
architects to have any lateral freedom to design a working
implementation with a low-cost path to verification. I don't see why
there's any fundamental reason that "scalable" approaches he's pointed to
can't be fit into our methodology directly? I've only glanced at the
papers, but it seems like it should be so.
--- Note sure how to condense it and incorporate -Murali

Reviewer-F:
1. Evidence of proofs being machine-checked:
We have submitted the source code of our Coq development as a supplement to our
paper in the submission website.

2. Figure 3, Tags in Figures 5 & 6:
Both these points are valid and we could indicate these points explicitly.

3. Figure 6: The third case:
Reviewer-F's assertion is not true. The "upd" function merely determines if the
"real" state of the system is as dictated by the state (s,pc) or if it should
be modified to take into account the responses from the memory system. If the
processor is not waiting for the memory system response, then the state is
exactly (s,pc). We made an optimization in the fourth case where we make the
real state H whenever we know that the next instruction we decode is H; instead
the state can be left as (s,pc) in the fourth case.

4. Figure 8: Tags for non-speculative loads:
Our invariant is that there can be only one outstanding non-speculative load,
though there can be several outstanding speculative loads along with the
outstanding non-speculative load. Thus, anything without a tag is the
non-speculative load and there is only one of them, and anything with a tag is
a speculative load.

Reviewer-E's related work summary:

FM methodology: The methodology used for proving that the memory system
realizes cache coherence is not new. A theorem-proving approach to verifying
memory systems has been tacit in many publications. Researchers preferred model
checking, as opposed to pushing through theorem-proving based proofs largely
because of the daunting complexity of realistic memory system protocols. One
can find several approaches that address specific aspects of memory system
verification in, say, Wim Hesselink's work :
http://www.informatik.uni-trier.de/~ley/pers/hd/h/Hesselink:Wim_H=.

@@@ Murali response
Lots of PVS-based proofs for concurrent algorithms, mutual exclusion, etc...
(Needs to be read further, but it's just a bunch of mechanized proofs for
a lot of concurrent algorithms)
***

There has of course been recent trends to simplify protocols thru the DeNoVo
project:
http://scholar.google.com/citations?view_op=view_citation&hl=en&user=ufGYoL0AAAAJ&citation_for_view=ufGYoL0AAAAJ:8k81kl-MbHgC
The authors have not cited such related pieces of work and justified the
simplicity of their design.

@@@ Murali response
A new non-conventional cache coherence protocol that produces sequentially
consistent results for data-rate free programs. There's no formal proof that
this protocol is correct...
(Needs to be read further, but doesn't seem relevant)
***

The authors claim that they are achieving their proof for an arbitrary number
of processors. They do not mention anything about the invariants involved in
their parametric proofs. This leads me to suspect that these invariants are
pretty straightforward.  If one harks back to work addressing real protocols,
the Chou, Mannava, Park work
(http://scholar.google.com/citations?view_op=view_citation&hl=en&user=ufGYoL0AAAAJ&citation_for_view=ufGYoL0AAAAJ:qjMakFHDy7sC)
shows how exceedingly hard it is to discover these invariants. In fact, the CMP
approach is an incremental CEGAR-style method for discovering these invariants.
Stating these invariants explicitly would be virtually impossible - even for
FLASH-like protocols.  In fact, a whole series of papers have tried to address
invariant discovery in this space : http://dl.acm.org/citation.cfm?id=1517434

@@@ Murali response
Chou, Mannava, Park [CMP-method] approach uses three paradigms: a) symmetry of
cache coherence protocols b) abstraction of several nodes which behaves as just
another node and c) counter-example guided abtraction refinement (CEGAR) for
refining the abstraction's properties. Using CEGAR, they get invariants for the
rest of the system while concentrating on a finite number of nodes (2).  Each
refinement will not restrict the behavior, but will help towards proving a
safety property. Specifically, let P be an invariant for the abstract
transition system, and we should prove that it satisfies safety property X. We
start out with a P that doesn't satisfy X. Then we add constraints on P, while
permitting the same transitions. Eventually P will satisfy X. The manual
portion of this procedure is in strengthening P.
***

@@@ Murali response
The second paper tries to formulate better invariants in the first place using
the notion of "flows". It says what other messages can interfere when the other
message is being processed...
(Needs to be read further -- I don't understand "flows" clearly; can't say
anything more before I understand that properly)
***

For hierarchical protocols, this work has been extended in e.g.,
http://link.springer.com/article/10.1007%2Fs10703-010-0092-y

@@@ Murali response
Same approach as [CMP-method] to hierarchical protocols.
***
